---
title: "sta130proj"
output:
  pdf_document: default
  html_document: default
date: "2023-03-18"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## R Markdown


```{r cars}
library(rpart)
library(tidyverse)
library(partykit)
```

```{r}
# load basic data
library(rhdf5)
header <- h5ls("STA130_APOGEE.h5")
header
```

```{r}
et = "STA130_APOGEE.h5" %>% h5read("ca_uh")

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

*Abstract*
*This study aims to investigate the relationship between the effective temperature of positive and negative Ca/H stars. Specifically, the research question is whether the effective temperature is similar or different for these two types of stars. To test this, two hypotheses were formulated: the null hypothesis posits that the effective temperatures of positive Ca/H and negative Ca/H stars are generally similar, while the alternative hypothesis suggests that they are different.*

*To examine these hypotheses, two visualizations were created: a histogram and a boxplot were used to observe the distribution of the data. In addition, a two-sided hypothesis test was conducted to determine whether the null hypothesis should be rejected or not.*


*Method*
*We choosing the method of Two Sample hypothesis Test(permutation test), to analysis and study if the effective temperature of stellar with a positive Ca/H be same as the stellar with a negative Ca/H values.*
*$H_{0}: \text{ch}{1} = \text{ch}{2}$*
*$H_{0}$ state the stellar with positive [Ca/H]($\text{ch}{1}$) generally have same effective temperatures to the stellar with negative [Ca/H] values($\text{ch}{2}$)*
*$H_{1}: \text{ch}{1}  \neq {ch}{2}$*
*$H_{0}$ state the stellar with positive [Ca/H]($\text{ch}{1}$) generally have different effective temperatures to the stellar with negative [Ca/H] values($\text{ch}{2}$)*
*The significant level was set to be 0.05, so if the p-value is smaller than 0.05, we will reject $H_{0}$, if it is larger than 0.05, we will support $H_{1}$.*
*Firstly, we calculate the observed test statistic of difference in mean value of the temperature between stars with positive Ca/H and negative Ca/, which is about 22 degree celsius. Then we perform our simulation with 1000 times of loop iterations, each time shuffling the data ramdomly into two groups and calculate the difference mean of the negative and positive Ca/H's temperature without repalcement, then calculated the difference of the stimulated mean temperature, stored it in the vector simulated_values.*
*Then we calculated the p-value using the observed test statistic and simulated_values(the sampling distribution by assuming $H_{0}$ is true), and compare with the significant level(in this case is 0.05).*


*P-value result interpretation*
*Our calculated p value is 0, which is very small. Our calculated p-value is not necessary the same as the actual p-value, but this implies that a 22 degree celsius difference in mean effective temperature did not happened within 1000 times of random sampling. Thus, the actual p-value must be very small. Since our p-value is definitely smaller than the significant level of 0.05, we reject the $H_{0}$ state the stellar with positive [Ca/H]($\text{ch}{1}$) generally have same effective temperatures to the stellar with negative [Ca/H] values($\text{ch}{2}$), which means though the permutation test, we support the $\text{ch}{1}$, which the temperature of stellar with positive [Ca/H]($\text{ch}{1}$) generally have different effective temperatures to the stellar with negative [Ca/H] values($\text{ch}{2}$).*

```{r}

# this is the first visualization of our first research quesiton
cah <- "STA130_APOGEE.h5" %>% h5read("ca_h")
et = "STA130_APOGEE.h5" %>% h5read("teff")
cah_et = data.frame(cah, et)
cah_et_tibble = cah_et %>% as_tibble()
cah_abundance_tibble <- cah_et_tibble %>% mutate(abundance = case_when(cah < 0 ~ 0, cah >= 0 ~ 1))

cahlow = cah_et_tibble %>% filter(cah <= 0)
cahhigh = cah_et_tibble %>% filter(cah > 0)
mean0 = mean(cahlow$et)
mean1 = mean(cahhigh$et)

test_stats = mean0 - mean1
simulated_values <- 1:1000

for(i in 1:1000){
  simdata <- cah_abundance_tibble %>% mutate(abundance = sample(abundance, replace = FALSE))
  sim_value <- simdata %>% group_by(abundance) %>%
  summarise(mean = mean(et), .groups="drop") %>%
  summarise(value = diff(mean))

  simulated_values[i] <- as.numeric(sim_value)
  
}
sim_tibble = simulated_values %>% as_tibble()
sim_tibble
ggplot(data = sim_tibble, aes(x = value)) + geom_histogram() + labs(x="Difference of sample mean effctive tempreature of postive cah and negative cah stellar")

p = sum(abs(simulated_values) >= abs(test_stats))/1000
tibble(pvalue-question1=p)
```

```{r}
cah_mutated =cah_et_tibble %>% mutate(boolean = case_when(cah > 0 ~ "cah>0", cah <= 0 ~ "cah<=0"))%>%group_by(boolean)
cah_mutated %>%
  ggplot(aes(x = boolean, y = et)) +
  geom_boxplot(color = "black", fill = "grey") +
  labs(x="cah range", y="temperature(celcius degree)")
```


```{r}


Al_h <- "STA130_APOGEE.h5" %>% h5read("al_h")
mg_h <- "STA130_APOGEE.h5" %>% h5read("mg_h")
fe_h <- "STA130_APOGEE.h5" %>% h5read("fe_h")
heavyele_et = data.frame(Al_h, mg_h, fe_h, cah, et)
heavyele_et_tibble = heavyele_et %>% as_tibble()

heavyele_abundance_tibble <- heavyele_et_tibble %>% mutate(abundance = case_when(Al_h + mg_h + fe_h + cah < 0 ~ 0, Al_h + mg_h + fe_h + cah >= 0 ~ 1))
heavyele_abundance_tibble
heavyelelow = heavyele_abundance_tibble %>% filter(abundance == 0)
heavyelehigh = heavyele_abundance_tibble %>% filter(abundance > 0)
mean2 = mean(heavyelelow$et)
mean3 = mean(heavyelehigh$et)
test_stat2 = mean3 - mean2


#permutation test start here
simulated_values2 <- 1:100

for(i in 1:100){
  simdata <- heavyele_abundance_tibble %>% mutate(abundance = sample(abundance, replace = FALSE))
  sim_value <- simdata %>% group_by(abundance) %>%
  summarise(mean = mean(et), .groups="drop") %>%
  summarise(value = diff(mean))

  simulated_values2[i] <- as.numeric(sim_value)
  
}
sim_tibble2 = simulated_values2 %>% as_tibble()

ggplot(data = sim_tibble2, aes(x = value)) + geom_histogram()
test_stat2
sim_tibble2
p = sum(abs(simulated_values2) >= abs(test_stat2))/100
p

```


```{r}
# this is the linear regression analyze of our second research problem

et = "STA130_APOGEE.h5" %>% h5read("teff")
loggra = "STA130_APOGEE.h5" %>% h5read("logg")
df <- data.frame(loggra, et)
model <- lm(et ~ loggra, data = df) %>% summary
ggplot(df, aes(x = loggra, y = et)) +
  geom_point() +
  geom_smooth(method = "lm") + labs(x="log base 10 gravity of stellar", y="effctive temperature of stellar(celcius degree)")

model
correnum=df %>% summarise(correlation = 
                              cor(et, loggra))
correnum
```
```{r}
mean_gra = mean(loggra)
etgra_tibble <- df %>% as_tibble()
lowetgroup <- etgra_tibble %>% filter(loggra >= mean_gra)
highetgroup <-etgra_tibble %>% filter(loggra < mean_gra)
etgra_tibble_mutate <- etgra_tibble %>% mutate(gragroup = case_when(loggra >= mean_gra ~1, loggra<mean_gra~0))
test_sta_r2 = abs(mean(lowetgroup$et) - mean(highetgroup$et))

simulated_values_r2 <- 1:1000

for(i in 1:1000){
  simdata <- etgra_tibble_mutate %>% mutate(et = sample(et, replace = FALSE))
  sim_value <- simdata %>% group_by(gragroup) %>%
  summarise(mean = mean(et), .groups="drop") %>%
  summarise(value = diff(mean))

  simulated_values_r2[i] <- as.numeric(sim_value)

}
sim_tibble_r2 = simulated_values_r2 %>% as_tibble()
sim_tibble_r2
ggplot(data = sim_tibble_r2, aes(x = value)) + geom_histogram()

p = sum(abs(simulated_values_r2) >= abs(test_sta_r2))/1000
p
```

```{r}
test_sta_r2
```



```{r}
# this is the analyze of question 3
#load data
feh <- "STA130_APOGEE.h5" %>% h5read("fe_h") %>% as_tibble() 
cah_tibble <- "STA130_APOGEE.h5" %>% h5read("ca_h") %>% as_tibble() 
alh <- "STA130_APOGEE.h5" %>% h5read("al_h") %>% as_tibble() 
mgh <- "STA130_APOGEE.h5" %>% h5read("mg_h") %>% as_tibble() 
```

```{r}

#this is the new linear regression analyze and visualization of heavy elements respectively

df4 = data.frame(loggra, cah = cah_tibble$value, alh = alh$value, mgh = mgh$value, feh = feh$value)

lm(loggra ~ mgh, data = df4) %>% summary
lm(loggra ~ feh, data = df4) %>% summary
lm(loggra ~ alh, data = df4) %>% summary
lm(loggra ~ cah, data = df4) %>% summary

```
```{r}
#four scatter plot visualization
df4 %>% ggplot(aes(x = mgh, y = loggra)) +
  geom_point() +
  geom_smooth(method = "lm")

df4 %>% ggplot(aes(x = feh, y = loggra)) +
  geom_point() +
  geom_smooth(method = "lm")

df4 %>% ggplot(aes(x = alh, y = loggra)) +
  geom_point() +
  geom_smooth(method = "lm")

df4 %>% ggplot(aes(x = cah, y = loggra)) +
  geom_point() +
  geom_smooth(method = "lm")

```



```{r}
df4 %>% summarise(correlationfeh =
                              cor(loggra, feh))
df4 %>% summarise(correlationcah = cor(loggra, cah))
df4 %>% summarise(correlationmgh =
                              cor(loggra, mgh))
df4 %>% summarise(correlationalh =
                              cor(loggra, alh))

```


```{r}
set.seed(130)  # required to ensure reproducibility
mean_gra2 = mean(df4$loggra)
new_df4 <- df4 %>% mutate(gra_catogory = ifelse(loggra >= mean_gra2, 'high', 'low')) %>% mutate(sum_heavyelement = cah + mgh + alh + feh)
tree1 <- rpart(gra_catogory ~ sum_heavyelement, data = new_df4)
plot(as.party(tree1), type="simple", gp=gpar(cex=0.5))
```

```{r}
tree1pre = predict(tree1, newdata = new_df4) %>% as_tibble() %>% mutate(prediction = ifelse(high >= 0.5, 'predict high','predict low'))
cm = table(tree1pre$prediction, new_df4$gra_catogory)
cm
```

```{r}
precision <- cm[1, 1] / sum(cm[1,])
recall <- cm[1, 1] / sum(cm[,1])
accuracy <- sum(diag(cm)) / sum(cm)
precision
recall
accuracy
```


