---
title: "sta130proj"
output:
  pdf_document: default
  html_document: default
date: "2023-03-18"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## R Markdown


```{r cars}

library(tidyverse)
```

```{r}
# load basic data
library(rhdf5)
header <- h5ls("STA130_APOGEE.h5")
header
```

```{r}
glimpse(header)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.



*Method*
*We choosing the method of Two Sample hypothesis Test(approximation test) which is two sided hypothesis test, to analysis and study if the effective temperature of stellar with a positive Ca/H be same as the stellar with a negative Ca/H.  temperatures similar to the stellar with negative [Ca/H] values.*
*$H_{0}: \text{ch}{1} = \text{ch}{2}$*
*$H_{0}$ state the stellar with positive [Ca/H]($\text{ch}{1}$) generally have same effective temperatures to the stellar with negative [Ca/H] values($\text{ch}{2}$)*
*$H_{1}: \text{ch}{1}  \neq {ch}{2}$*
*$H_{0}$ state the stellar with positive [Ca/H]($\text{ch}{1}$) generally have different effective temperatures to the stellar with negative [Ca/H] values($\text{ch}{2}$)*
*The significant level was set to be 0.05, so if the p-value is smaller than 0.05, we will reject $H_{0}$, if it is larger than 0.05, we will support $H_{1}$.*
*Firstly, we calculate the observed test statistic of mean value of the temperature positive Ca/H and negative Ca/H and their difference(is 22 Celsius degree). Then we do 100 number of loop, each time shuffling the "Ca/H" variable and calculate the mean of the negative and positive Ca/H's temperature by bootstrap method, then calculated the difference of the stimulated mean temperature, stored it in simulated_values[i].*
*Then we calculated the p-value using the observed test statistic and simulated_values(the sampling distribution by assuming $H_{0}$ is true), and compare with the significant level(in this case is 0.05).*


*P-value result interpretation*
*We get a p-value of 0 in this case(since we have around 90000 data so it is very likely to get a p-value of 0).And 0 is smaller than the significant level of 0.05 so we reject the $H_{0}$ state the stellar with positive [Ca/H]($\text{ch}{1}$) generally have same effective temperatures to the stellar with negative [Ca/H] values($\text{ch}{2}$), which means though the Two sample hypothesis test, we support the $\text{ch}{1}$, which the temperature of stellar with positive [Ca/H]($\text{ch}{1}$) generally have different effective temperatures to the stellar with negative [Ca/H] values($\text{ch}{2}$).*



```{r}

# this is the first visualization of our first research quesiton
df0 <- data.frame(cah_abundance, et)#two variables are defined below
lm(et ~ abundance, data = df0) %>% summary

ggplot(df0, aes(x = abundance, y = et)) +
  geom_point() +
  geom_smooth(method = "lm")
```


```{r}


# this is the linear regression analyze of our second research problem
et = "STA130_APOGEE.h5" %>% h5read("teff")
loggra = "STA130_APOGEE.h5" %>% h5read("logg")
df <- data.frame(loggra, et)
model <- lm(et ~ loggra, data = df) %>% summary
ggplot(df, aes(x = loggra, y = et)) +
  geom_point() +
  geom_smooth(method = "lm")

model
```

```{r}
# this is multilinear regression analyze of question 3
feh <- "STA130_APOGEE.h5" %>% h5read("fe_h") %>% as_tibble() 
feh_abundance <- feh %>% mutate(abundance = case_when(value < 0 ~ 0, value >= 0 ~ 1))

cah <- "STA130_APOGEE.h5" %>% h5read("ca_h") %>% as_tibble() 
cah_abundance <- cah %>% mutate(abundance = case_when(value < 0 ~ 0, value >= 0 ~ 1))

alh <- "STA130_APOGEE.h5" %>% h5read("al_h") %>% as_tibble() 
alh_abundance <- alh %>% mutate(abundance = case_when(value < 0 ~ 0, value >= 0 ~ 1))

mgh <- "STA130_APOGEE.h5" %>% h5read("mg_h") %>% as_tibble() 
mgh_abundance <- mgh %>% mutate(abundance = case_when(value < 0 ~ 0, value >= 0 ~ 1))

df2 <- data.frame(loggra, cah = cah_abundance, alh = alh_abundance, mgh = mgh_abundance, feh = feh_abundance)
df2

model3 <- lm(loggra ~ cah.abundance + alh.abundance +mgh.abundance +  feh.abundance, data = df2) %>% summary
model3


df3 <- data.frame(loggra, ca_al_abundance_interact = cah$value + alh$value, cah = cah$value, alh = alh$value, mgh = mgh$value, feh = feh$value)

df3 %>% ggplot(aes(x = loggra, y = ca_al_abundance_interact)) +
  geom_point() +
  geom_smooth(method = "lm")


```

```{r}

#this is the new linear regression analyze and visualization of heavy elements respectively

df4 = data.frame(loggra, cah = cah$value, alh = alh$value, mgh = mgh$value, feh = feh$value)

lm(loggra ~ mgh, data = df4) %>% summary
lm(loggra ~ feh, data = df4) %>% summary
lm(loggra ~ alh, data = df4) %>% summary
lm(loggra ~ cah, data = df4) %>% summary
df4 %>% ggplot(aes(x = loggra, y = mgh)) +
  geom_point() +
  geom_smooth(method = "lm")

df4 %>% ggplot(aes(x = loggra, y = feh)) +
  geom_point() +
  geom_smooth(method = "lm")

df4 %>% ggplot(aes(x = loggra, y = alh)) +
  geom_point() +
  geom_smooth(method = "lm")

df4 %>% ggplot(aes(x = loggra, y = cah)) +
  geom_point() +
  geom_smooth(method = "lm")

```

```{r}
df4 %>% summarise(correlationfeh = 
                              cor(loggra, feh))
df4 %>% summarise(correlationcah = 
                              cor(loggra, cah))
df4 %>% summarise(correlationmgh = 
                              cor(loggra, mgh))
df4 %>% summarise(correlationalh = 
                              cor(loggra, alh))
```
```{r}
summary(lm(feh ~ loggra, data = df4))
summary(lm(cah ~ loggra, data = df4))
summary(lm(mgh ~ loggra, data = df4))
summary(lm(alh ~ loggra, data = df4))
```
